On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted

### 2016-03-06 MS.tex
typo
Add Ian's comments

In this paper, we consider five prominent traits (specific leaf area = SLA, wood or stem tissue density, maximum leaf photosynthetic capacity per unit area = Aarea, maximum or asymptotic height, and seed mass) and we investigate:
\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item Do the observed relationships conform to predictions derived from a recently developed mechanistic model \citep{Falster:2011ii} (see ``Theoretical expectations'')?
\end{enumerate}
To address these questions we used a meta-analytic approach, compiling data from studies reporting correlation coefficients between traits and growth. While most studies investigating the link between traits and growth have not primarily been concerned with the effects of plant size, they have generally reported trait-growth correlations for identifiable ontogenetic stages (i.e. seedling, sapling, and adult). We therefore used ontogenetic stage as a surrogate for plant size. We focused on tree and woody species since the mechanistic model used to assess our predictions is most applicable to these growth forms.
Using a meta-analysis approach, we were able to assess our hypotheses across a much broader number of species and vegetation types than would have been possible in an experiential study. We anticipated a sizeable literature on traits and growth to be available but we were uncertain about the representation of all trait-stage combinations. An additional goal of our study was therefore to identify gaps in the knowledge needed to improve our understanding of the link between functional traits and plant growth.

@@

Assuming plants maintain a similar cross section of conducting tissue, cheaper stem construction (low wood density or stem density) will increase the marginal rate of leaf deployment per total mass increment ($\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$), and consequently growth rate in height and diameter (eqs. \ref{eq:eq2}-\ref{eq:eq4}; Table \ref{tab:trade-offs}). Disadvantages associated with cheap stem construction take the form of higher mortality and so are not reflected in growth equations. Under this scenario, wood density is negatively correlated to growth rate across all ontogenetic stages. Because the relative expenditure on stem compared to leaf increases as plants grow larger, the relationship between wood density and growth may also strengthen in larger plants. A similar shift is expected for growth in mass, in either absolute or relative terms. Note in the case of mass growth, the effect of cheaper stem construction is realised not in the $\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$ term , but via greater total leaf area at a given mass, i.e. the accumulated effect of past reductions in  $\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$.

@@

As plants approach maximum height they increasingly divert energy away from growth and into reproduction (see \citealt{Thomas:1996do,Thomas-2011, Wenk:2015jz}). This effect occurs in the term $\frac{\ud M_{\rm a} }{\ud B}$ (the allocation to vegetative growth) in eqs. (\ref{eq:eq2}-\ref{eq:eq3}). For seedlings, where all individuals are investing solely in growth, we expect no influence of maximum height on growth (i.e. $\frac{\ud M_{\rm a} }{\ud B} =1$). But once some species begin to produce seed, then species closer to their maximum height value are expected to be allocating more to seed production and less to vegetative growth. Hence when comparing across plants at a given size, species with smaller maximum height are expected to show slower vegetative growth rates. Similar correlations are expected no matter what measure of growth is used (Table \ref{tab:trade-offs}).

@@

For maximum height, the expectation was for a correlation with growth to be present for large plants but absent for smaller sizes, the mechanism being increased relative allocation to reproduction as maximum height is approached. Here, maximum height was positively correlated with growth at adult stage ($\bar{r}= 0.38 \pm 0.05$), not correlated at sapling stage (Fig \ref{fig:fig3}c, CIs overlapping zero) and no data were recorded for seedling stage.
While these results were consistent with our expectation that maximum height would be correlated with growth at adult stage, they did not differ significantly across ontogenetic stages (model fit not improved when stage was incorporated, see log likelihood ratio tests in Fig. \ref{fig:fig3}c, and CIs overlapping between sapling and adult stages). Due to small numbers of replicate studies for sapling stage and no replicates for seedling stage, we could not conclude on the absence of correlation between maximum height and growth at small plant size.

@@

Results from the entire dataset reinforced our conclusions for both SLA and wood density, and did not alter them for the others traits. For SLA and wood density, although the number of replicates doubled by using the entire dataset, effect sizes and overall results were similar to those reported above using the conservative crieteria (Fig. \ref{fig:fig3})). For seed mass and Aarea, the additional replicates in the entire dataset allowed us to be more confident of the effect size for seedling stage (Fig \ref{fig:fig3}d, e). Comparison of Aarea effect between the conservative and relaxed criteria (Fig. \ref{fig:fig3}d), indicated little effect of stress (here mainly low light) on the correlation with relative growth rate at seedling stage as the correlation was still positive when 8 correlations measured under low light conditions were added. For maximum height, there were no additional replicates in the entire dataset; we still cannot conclude undoubtedly that the correlation is absent for smaller sizes.

@@

The only correlation that we hypothesized would change if absolute growth rate was used instead of relative growth rate, was for eed mass. Overall, the models fit did not improve when growth measurement was incorporated (see log likelihood ratio tests in Fig. \ref{fig:fig4}). For a majority of traits, effect sizes were generally consistent between relative growth rate and absolute growth rate when both measurements were available (e.g. effect size for relative and absolute growth rate of Hmax and Wood density at adult stage in Fig. \ref{fig:fig4} b and c). However, for seed mass and wood density at seedling stage, the direction of effect sizes changed when correlations were measured with abolute versus relative growth rate. For seed mass, effect size at seedling stage with relative growth rate was positive and shifted to non significant with absolute growth rate (Fig \ref{fig:fig4}d; CI overlapping zero). While these results were consistent with expectations, our ability to conclude was limited due to low sample size for correlation with absolute growth rate (i.e CIs overlapping, Fig. \ref{fig:fig4}d). For wood density, some differences occurred according to the growth measurement used, but they were not easily attributed to any single cause. In particular, the correlation at juvenile stage shifted from negative with relative growth rate to non significant with absolute growth rate (i.e CI included zero, Fig. \ref{fig:fig4}b). Yet, the correlation coefficients recorded for absolute growth rate came from \citet{Augspurger:1984ct}, and unlike others they used wood density values from adult plants together with growth from juveniles. So it was not possible to disentangle the effect of growth measurement from the effect of trait origin.

>>>>

In this paper, we consider five prominent traits (specific leaf area = SLA, wood or stem tissue density, maximum leaf photosynthetic capacity per unit area = Aarea, maximum or asymptotic height, and seed mass) and we investigate:
\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item Do the observed relationships conform to predictions derived from a recently developed mechanistic model \citep{Falster:2011ii} (see ``Theoretical expectations'')?
\end{enumerate}
To address these questions we used a meta-analytic approach, compiling data from studies reporting correlation coefficients between traits and growth. While most studies investigating the link between traits and growth have not primarily been concerned with the effects of plant size, they have generally reported trait-growth correlations for identifiable ontogenetic stages (i.e. seedling, sapling, and adult). We therefore used ontogenetic stage as a surrogate for plant size. We focused on tree and woody species since the mechanistic model used to assess our predictions is most applicable to these growth forms.
Using a meta-analysis approach, we were able to assess our hypotheses across a much broader number of species and vegetation types than would have been possible in an experimental study. We anticipated a sizeable literature on traits and growth to be available but we were uncertain about the representation of all trait-stage combinations. An additional goal of our study was therefore to identify gaps in the knowledge needed to improve our understanding of the link between functional traits and plant growth.

@@

Assuming plants with a similar cross section of conducting tissue, cheaper stem construction (low wood density or stem density) will increase the marginal rate of leaf deployment per total mass increment ($\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$), and consequently growth rate in height and diameter (eqs. \ref{eq:eq2}-\ref{eq:eq4}; Table \ref{tab:trade-offs}). Disadvantages associated with cheap stem construction take the form of higher mortality and so are not reflected in growth equations. Under this scenario, wood density is negatively correlated to growth rate across all ontogenetic stages. Because the relative expenditure on stem compared to leaf increases as plants grow larger, the relationship between wood density and growth may also strengthen in larger plants. A similar shift is expected for growth in mass, in either absolute or relative terms. Note in the case of mass growth, the effect of cheaper stem construction is realised not in the $\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$ term , but via greater total leaf area at a given mass, i.e. the accumulated effect of past reductions in  $\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$.

@@

As plants approach maximum height they increasingly divert energy away from growth and into reproduction (see \citealt{Thomas:1996do,Thomas-2011, Wenk:2015jz}). This effect occurs in the term $\frac{\ud M_{\rm a} }{\ud B}$ (the allocation to vegetative growth) in eqs. (\ref{eq:eq2}-\ref{eq:eq3}). For seedlings, where individuals are investing solely in growth, we expect no influence of maximum height on growth (i.e. $\frac{\ud M_{\rm a} }{\ud B} =1$). But once some species begin to produce seed, then species closer to their maximum height value are expected to be allocating more to seed production and less to vegetative growth. Hence when comparing across plants at a given size, species with smaller maximum height are expected to show slower vegetative growth rates. Similar correlations are expected no matter what measure of growth is used (Table \ref{tab:trade-offs}).

@@

For maximum height, the expectation was for a correlation with growth to be present for large plants but absent for smaller sizes, the mechanism being increased relative allocation to reproduction as maximum height is approached. Here, maximum height was positively correlated with growth at adult stage ($\bar{r}= 0.38 \pm 0.05$), not significantly correlated at sapling stage (Fig \ref{fig:fig3}c, CIs overlapping zero) and no data were recorded for seedling stage.
While these results were consistent with our expectation that maximum height would be correlated with growth at adult stage, they did not differ significantly across ontogenetic stages (model fit not improved when stage was incorporated, see log likelihood ratio tests in Fig. \ref{fig:fig3}c, and CIs overlapping between sapling and adult stages). Due to small numbers of replicate studies for sapling stage and no replicates for seedling stage, we could not conclude on the absence of correlation between maximum height and growth at small plant size.

@@

Results from the entire dataset reinforced our conclusions for both SLA and wood density, and did not alter them for the others traits. For SLA and wood density, although the number of replicates doubled by using the entire dataset, effect sizes and overall results were similar to those reported above using the conservative crieteria (Fig. \ref{fig:fig3}). For seed mass and Aarea, the additional replicates in the entire dataset allowed us to be more confident of the effect size for seedling stage (Fig \ref{fig:fig3}d, e). Comparison of Aarea effect between the conservative and relaxed criteria (Fig. \ref{fig:fig3}d), indicated little effect of stress (here mainly low light) on the correlation with relative growth rate at seedling stage as the correlation was still positive when 8 correlations measured under low light conditions were added. For maximum height, there were no additional replicates in the entire dataset; we still cannot conclude undoubtedly that the correlation is absent for smaller sizes.

@@

The only correlation that we hypothesized would change if absolute growth rate was used instead of relative growth rate, was for seed mass. Overall, the models fit did not improve when growth measurement was incorporated (see log likelihood ratio tests in Fig. \ref{fig:fig4}). For a majority of traits, effect sizes were generally consistent between relative growth rate and absolute growth rate when both measurements were available (e.g. effect size for relative and absolute growth rate of Hmax and Wood density at adult stage in Fig. \ref{fig:fig4} b and c). However, for seed mass and wood density at seedling stage, the direction of effect sizes changed when correlations were measured with abolute versus relative growth rate. For seed mass, effect size at seedling stage with relative growth rate was positive and shifted to non significant with absolute growth rate (Fig \ref{fig:fig4}d; CI overlapping zero). While these results were consistent with expectations, our ability to conclude was limited due to low sample size for correlation with absolute growth rate (i.e CIs overlapping, Fig. \ref{fig:fig4}d). For wood density, some differences occurred according to the growth measurement used, but they were not easily attributed to any single cause. In particular, the correlation at juvenile stage shifted from negative with relative growth rate to non significant with absolute growth rate (i.e CI included zero, Fig. \ref{fig:fig4}b). Yet, the correlation coefficients recorded for absolute growth rate came from \citet{Augspurger:1984ct}, and unlike others they used wood density values from adult plants together with growth from juveniles. So it was not possible to disentangle the effect of growth measurement from the effect of trait origin.
